Communication complexity: Chapter 5 \\ Information complexity

### 2019-02-08 CS860-CommunicationComplexity/assignments/Chapter5.tex

\begin{theorem}
Entropy satisfies the following properties:
\begin{description}
\item[Boundedness] For every $Z$ over a finite set $\calZ$, $0 \le H(Z) \le \log_2 |\calZ|$.
\item[Chain rule] $H(Z_1\,Z_2\cdots Z_k) = H(Z_1) + H(Z_2 \mid Z_1) + \cdots + H(Z_k \mid Z_1\cdots Z_{k-1})$.
\item[Subadditivity] $H(Z \mid Z') \le H(Z)$ and $H(Z\,Z') \le H(Z) + H(Z')$.
\end{description}
Mutual information satisfies the following properties:
\begin{description}
\item[Boundedness] $0 \le I(Z ; Z') \le \min\{ H(Z), H(Z') \}$.
\item[Chain rule] $I(Z_1\,Z_2 ; W) = I(Z_1 ; W) + H(Z_2 ; W \mid Z_1)$.
\item[Data processing inequality] Whenever $Z' = f(Z)$ is determined by $Z$, $I(Z' ; W) \le I(Z ; W)$.
\end{description}

@@

\begin{definition}[Internal information cost]
The \emph{internal information cost} of a randomized protocol $\Pi$ over the distribution $\mu$ on $\calX \times \calY$ is
\[
\icostint_\mu(\Pi) = I\big( X ; \Gamma_{X,Y}^\Pi \, Y\big) + I\big( Y ; \Gamma_{X,Y}^\Pi \, X\big).
\]
where $(X,Y) \sim \mu$.
\end{definition}

@@

\begin{theorem}
For every function $f : \calX \times \calY \to \{0,1\}$, every distribution $\mu$ on $\calX \times \calY$, and every $0 \le \epsilon \le \frac12$,
\[
\ICint_{\mu,\epsilon}(f) \le \ICext_{\mu,\epsilon}(f).
\]
\end{theorem}

@@

Given a function $f : \calX \times \calY \to \{0,1\}$, define $f^k : \calX^k \times \calY^k \to \{0,1\}^n$ to be the function where
\[
f^n(x^{(1)},\ldots,x^{(k)},y^{(1)},\ldots,y^{(k)}) = \big( f(x^{(1)},y^{(1)}), \ldots f(x^{(k)}, y^{(k)})\big).
\]
This corresponds to the setting where Alice and Bob must compute the value of $f$ on $n$ different pairs of inputs instead of just $1$.

@@

One of the main advantages of working with internal information complexity is that in this setting every function \emph{does} satisfy the intuition that computing $k$ copies of a function is exactly $k$ times harder than computing a single copy of it.
\begin{theorem}
For every function $f : \calX \times \calY \to \{0,1\}$ and every $k \ge 1$,
\[
\ICint_{\mu,0}(f^k) = k \cdot \ICint_{\mu,0}(f).
\]
\end{theorem}

@@

\begin{lemma}
Let $\mu$ be the distribution on $\{0,1\} \times \{0,1\}$ defined by $\mu(0,0) = \mu(0,1) = \mu(1,0) = \frac13$.
For every $\epsilon > 0$, there is a constant $c_\epsilon > 0$ for which
\[
\ICint_{\mu,\epsilon}(\textsc{And}) = c_\epsilon.
\]
\end{lemma}

@@

\begin{proof}
\replacethistext{Enter your proof here.}
\end{proof}

>>>>

\begin{theorem}
Entropy satisfies the following properties:
\begin{description}
\item[Boundedness] For every $Z$ over a finite set $\calZ$, $0 \le H(Z) \le \log_2 |\calZ|$.
\item[Chain rule] $H(Z_1\,Z_2\cdots Z_k) = H(Z_1) + H(Z_2 \mid Z_1) + \cdots + H(Z_k \mid Z_1\cdots Z_{k-1})$.
\item[Subadditivity] $H(Z \mid Z') \le H(Z)$ and $H(Z\,Z') \le H(Z) + H(Z')$.
\end{description}
Mutual information satisfies the following properties:
\begin{description}
\item[Boundedness] $0 \le I(Z ; Z') \le \min\{ H(Z), H(Z') \}$.
\item[Chain rule] $I(Z_1\,Z_2 ; W) = I(Z_1 ; W) + I(Z_2 ; W \mid Z_1)$.
\item[Data processing inequality] Whenever $Z' = f(Z)$ is determined by $Z$, $I(Z' ; W) \le I(Z ; W)$.
\end{description}

@@

\begin{definition}[Internal information cost]
The \emph{internal information cost} of a randomized protocol $\Pi$ over the distribution $\mu$ on $\calX \times \calY$ is
\[
\icostint_\mu(\Pi) = I\big( X ; \Gamma_{X,Y}^\Pi\mid \, Y\big) + I\big( Y ; \Gamma_{X,Y}^\Pi \mid\, X\big).
\]
where $(X,Y) \sim \mu$.
\end{definition}

@@

\begin{theorem}\label{th:6}
For every function $f : \calX \times \calY \to \{0,1\}$, every distribution $\mu$ on $\calX \times \calY$, and every $0 \le \epsilon \le \frac12$,
\[
\ICint_{\mu,\epsilon}(f) \le \ICext_{\mu,\epsilon}(f).
\]
\end{theorem}

@@

Given a function $f : \calX \times \calY \to \{0,1\}$, define $f^k : \calX^k \times \calY^k \to \{0,1\}^k$ to be the function where
\[
f^k(x^{(1)},\ldots,x^{(k)},y^{(1)},\ldots,y^{(k)}) = \big( f(x^{(1)},y^{(1)}), \ldots f(x^{(k)}, y^{(k)})\big).
\]
This corresponds to the setting where Alice and Bob must compute the value of $f$ on $n$ different pairs of inputs instead of just $1$.

@@

One of the main advantages of working with internal information complexity is that in this setting every function \emph{does} satisfy the intuition that computing $k$ copies of a function is exactly $k$ times harder than computing a single copy of it.
\begin{theorem}\label{th:9}
For every function $f : \calX \times \calY \to \{0,1\}$ and every $k \ge 1$,
\[
\ICint_{\mu,0}(f^k) = k \cdot \ICint_{\mu,0}(f).
\]
\end{theorem}

@@

\begin{lemma}\label{lemma:1}
Let $\mu$ be the distribution on $\{0,1\} \times \{0,1\}$ defined by $\mu(0,0) = \mu(0,1) = \mu(1,0) = \frac13$.
For every $\epsilon > 0$, there is a constant $c_\epsilon > 0$ for which
\[
\ICint_{\mu,\epsilon}(\textsc{And}) = c_\epsilon.
\]
\end{lemma}

@@

\begin{proof}
Let $n=k$. Let $\pi$ be a communication protocol which computes $\Disj$ with error at most $\epsilon$, and achieves information cost
$$\ICint_{\mu,\epsilon}$$ . By Theorem \ref{th:2} and Theorem \ref{th:6},
$$R_\epsilon(\Disj) \geq \ICint_{\mu,\epsilon}(\Disj) = \icostint_{\mu,\epsilon}(\pi).$$
Notice that when we sample $(x^1,y^1), \dots, (x^k,y^k)$ according to $\mu$, then
$$\textsc{AND}^k((x^1,\dots,x^k),(y^1,\dots,y^k)) = \textsc{AND}(x^i,y^i) = 0$$
for any $i \in [k]$. Thus we can adapt any $\epsilon$-error protocol, $\rho$, for $\textsc{AND}^k$ to an $\epsilon$-error protocol, $\rho'$, for $\textsc{AND}$ in the same way that we did during the proof of Theorem \ref{th:9}, and consequently
$$\icostint_{\mu,\epsilon}(\rho') = \frac{1}{k}\icostint_{\mu,\epsilon}(\rho).$$
Actually our observation about $\textsc{AND}^k$ let's us say something stronger. We can use a protocol for $\Disj$ as a stand-in for a protocol for $\textsc{AND}^k$, since (when input is sampled according to $\mu$)
$$\Disj((x^1,\dots, x^k),(y^1,\dots,y^k)) = (x^1\wedge y^1) \lor\dots\lor(x^k\wedge y^k) = x^i\wedge y^i = \textsc{AND}^k((x^1,\dots, x^k),(y^1,\dots,y^k))$$
for any $i\in[k]$. Therefore we can adapt $\pi$ to a protocol for $\textsc{AND}$, $\pi'$ satisfying
$$\icostint_{\mu,\epsilon}(\pi') = \frac{1}{k}\icostint_{\mu,\epsilon}(\pi).$$
Thus using Lemma \ref{lemma:1},
$$\icostint_{\mu,\epsilon}(\pi) = k\icostint_{\mu,\epsilon}(\pi')\geq k\ICint_{\mu,\epsilon}(\textsc{AND}) = kc_\epsilon = \Omega(k).$$
Combining with our first observation we see that
$$R_\epsilon(\Disj) = \Omega(k)$$
as desired.
\end{proof}
